Computing the least commitment action models from observations of plan executions

### 2018-11-26 doc/ijcai-19-conformant/conformant-learning.tex
adding the classical planning and observation model to ijca19 paper on learning with conformant plannign

\def\year{2017}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{}
\setcounter{secnumdepth}{0}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage{comment}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}

\lstset{
  basicstyle=\ttfamily,
  mathescape
}

\usepackage{multicol}
\usepackage{arydshln}
\usetikzlibrary{calc,backgrounds,positioning,fit}

\newcommand{\tup}[1]{{\langle #1 \rangle}}

\newcommand{\pre}{\mathsf{pre}}     % precondition
\newcommand{\del}{\mathsf{del}}     % effect
\newcommand{\add}{\mathsf{add}}     % effect
\newcommand{\eff}{\mathsf{eff}}     % effect
\newcommand{\cond}{\mathsf{cond}}   % conditional effect
\newcommand{\true}{\mathsf{true}}   % true
\newcommand{\false}{\mathsf{false}} % false
\newcommand{\PE}{\mathrm{PE}}     % precondition
\newcommand{\strips}{\textsc{Strips}}     % precondition

@@

@@

\title{Model Recognition as Planning}

% Commented for blind submission
%\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
%{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
%{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
%{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
%{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
%\And Miquel Ram\'irez\\
%{\scriptsize School of Computing and Information Systems}\\
%{\scriptsize The University of Melbourne}\\
%{\scriptsize Melbourne, Victoria. Australia}\\
%{\scriptsize miquel.ramirez@nimelb.edu.au}}

\maketitle
\begin{abstract}
Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model that explains the observation. The paper formalizes this task and introduces a novel method that estimates the probability of a \strips\ model to produce an observation of a plan execution. This method builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and intermediate states in the observation. The effectiveness of the method is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a classical string classification task; (2) identification of the model that encodes a failure present in an observation; and (3) recognition of a robot navigation policy.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

A growing interest in addressing the task of recognizing the mission (plans and goals) of a planning agent from a set of given observations has recently emerged. Approaches based on rule-based systems, parsing, graph-covering or Bayesian nets have been proposed for the plan/goal recognition tasks~\cite{sukthankar14}. {\em Plan recognition as planning} is the model-based approach for this task that leverages the action model of the observed agent to compute its most likely goal and predict its future actions~\cite{ramirez2009plan,ramirez2012plan}. While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, in real-world applications developing in a particular context, approximate behavioral models of the involved agents (manually-introduced or automatically-learned) are already available, so there is rarely a need to learn a model from scratch. A more relevant and applicable task in these scenarios is thereby to predict the behavior of an agent from a set of given models.

This paper formalizes the novel task of {\em model recognition}, where the object to recognize is not a goal or a plan but the model that shapes the behavior of the observed agent. Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model in the set with the highest probability of producing the given observation. We posit this problem as a \emph{classification task} where classes are given as generative planning models and samples are sequences of observations. We accomplish this task with off-the-shelf classical planning algorithms that compute for each model (class) $\mathcal{M}$ in the set a model $\mathcal{M'}$ that explains the observation, and subsequently distances between the two models are measured. Two key contributions of our {\em model recognition as planning} approach are: (1) the ability of handling partial observations of plan executions that contain no actions and incomplete -- even empty- intermediate states; and (2) a thorough formalization that induces a probability distribution of a model $\mathcal{M}$ to explain an observed plan execution.

\begin{figure}
  \begin{scriptsize}
  \begin{center}
  \begin{tikzpicture}[scale=.4]
          \begin{scope}
            \draw (0, 0) grid (5, 5);
            \node[anchor=center] at (2.5, 1.5) {R};
            \node[anchor=center] at (0.5, 0.5) {};
            \draw[thick,style=dotted] (0.5,0.5) -- (3.5,0.5);
            \draw[thick,style=dotted] (3.5,0.5) -- (3.5,1.5);
            \draw[thick,style=dotted,->] (3.5,1.5) -- (2.7,1.5);
          \end{scope}
        \end{tikzpicture}
\hspace*{1cm}
	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
	  \node[state] (A)              {$q_0$};
	  \node[state] (B) [right of=A] {$q_1$};
	  \path
(A) edge [bend left, align=center] node {inc-y,\\dec-y} (B)
(B) edge [bend left, align=center] node {inc-y,\\dec-y} (A);
	\end{tikzpicture}
  \end{center}
  \end{scriptsize}
 \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid. {\em (Right)} Automata to control that the robot only increments its x-coordinate when $q_0$ holds (actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and switch the automata state).}
\label{fig:grid-example}
\end{figure}

To better illustrate {\em model recognition}, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}. According to this model, the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e. at even rows if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at odd rows if {\tt\small $q0$} holds initially). Besides this model, other models that shape different kinds of robot navigation can be defined within the same state variables (e.g. altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated). Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where state variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, {\em model recognition} aims to identify the robot action schemata that explains this observation.

We provide a formalization of {\em model recognition} that is applicable to any planning model but we restrict our attention to \strips\texttt{-}compilable models. It is well-known that diverse automata representations like finite state controllers, {\sc GOLOG} programs or reactive policies can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,segovia2017generating}. As for the evaluation, the effectiveness of our approach is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a proof-of-concept classification task; (2) a simulation of a non-deterministic \emph{blocksworld} under different degrees of observability; and (3) a navigation task where observations contain an unbounded number of unobservable states.

%: a classical string classification task; identification of the model that encodes a failure present in an observation; and recognition of a robot navigation policy.

%The paper also introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution. Our method is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.

\begin{figure}
  \begin{tiny}
  \begin{verbatim}
  (:action inc-x
    :parameters (?v1 ?v2)
    :precondition (and (xcoord ?v1) (next ?v1 ?v2) (q0))
    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))

  (:action dec-x
    :parameters (?v1 ?v2)
    :precondition (and (xcoord ?v1) (next ?v2 ?v1) (q1))
    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))

  (:action inc-y-even
    :parameters (?v1 ?v2)
    :precondition (and (ycoord ?v1) (next ?v1 ?v2) (q0))
    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
                 (not (q0)) (q1)))

  (:action inc-y-odd
    :parameters (?v1 ?v2)
    :precondition (and (ycoord ?v1) (next ?v1 ?v2) (q1))
    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
                 (not (q1)) (q0)))

  (:action dec-y-even
    :parameters (?v1 ?v2)
    :precondition (and (ycoord ?v1) (next ?v2 ?v1) (q0))
    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
                 (not (q0)) (q1)))

  (:action dec-y-odd
    :parameters (?v1 ?v2)
    :precondition (and (ycoord ?v1) (next ?v2 ?v1) (q1))
    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
                 (not (q1)) (q0)))
  \end{verbatim}
  \end{tiny}
 \caption{\small Example of a \strips\ action schemata (given in PDDL) for robot navigation in a $N\times N$ grid.}
\label{fig:model-example}
\end{figure}

\section{Background}
\label{sec:background}
This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.

\subsection{Classical planning with conditional effects}
$F$ is the set of {\em fluents} or {\em state variables} (propositional variables). A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. $L$ is a set of literals that represents a partial assignment of values to fluents, and $\mathcal{L}(F)$ is the set of all literals sets on $F$, i.e.~all partial assignments of values to fluents. A {\em state} $s$ is a full assignment of values to fluents. We explicitly include negative literals $\neg f$ in states and so $|s|=|F|$ and the size of the state space is $2^{|F|}$.

@@

Figure~\ref{fig:grid-example} illustrates a partial observation of a six-state trajectory \{{\tt\scriptsize<(xcoord 0)(ycoord 0)>, <(xcoord 1)(ycoord 0)>, <(xcoord 2)(ycoord 0)>, <(xcoord 3)(ycoord 0)>, <(xcoord 3)(ycoord 1)>, <(xcoord 2)(ycoord 1)>}\}. This observation only contains fluents of the predicates {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, corresponding to predicates {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unobservable in the six states.

\section{Model Recognition}
\label{sec:recognition}
The {\em model recognition} task is a tuple $\tup{P,M,\mathcal{O}}$ where:
\begin{itemize}
\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions. For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e. the functions $\rho$ and/or $\theta$ of $a$ are undefined.
\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_k\}$ is a set of {\em k different planning models} for the actions in $A[\cdot]$. A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$. Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
\end{itemize}

Model recognition can be understood as a {\em classification task} where each class is represented by a different planning model $\mathcal{M}\in M$, and the observed plan execution $\mathcal{O}(\tau)$ is the single example to classify. The planning model associated to each class acts as the corresponding {\em class prototype} and it summarizes any observation of a plan execution that could be synthesized with such model (i.e. the set of all the examples that belong to that class). We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to a given observation $\mathcal{O}(\tau)$. The {\em solution} to the model recognition task is then the subset of models in $M$ that maximizes this expression.

\begin{align}
argmax_{\mathcal{M}\in M} P(\mathcal{O}|\mathcal{M}) P(\mathcal{M}).
\end{align}

\subsection{Formulating the $P(\mathcal{O}|\mathcal{M})$ likelihood}

The $P(\mathcal{M})$ probability expresses whether one model is known to be a priori more likely than the others. If this probability is not given as input it is reasonable to assume that, {\em a priori}, all models are equiprobable. The challenge of our formulation to the model recognition task is the definition of $P(\mathcal{O}|\mathcal{M})$, the likelihood that expresses the probability of observing $\mathcal{O}(\tau)$ when $\mathcal{M}$ is the planning model.

\begin{figure}
  \begin{scriptsize}
  \begin{center}
	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
	  \node[state] (A)              {$\mathcal{M}$};
	  \node[state] (B) [right of=A] {$\mathcal{M}'$};
	  \node[state] (C) [right of=B] {$\tau$};
	  \node[state] (D) [right of=C] {$\mathcal{O}$};
	  \path
(A) edge  node {} (B)
(B) edge  node {} (C)
(C) edge  node {} (D)
;
	\end{tikzpicture}
  \end{center}
  \end{scriptsize}
 \caption{\small {\em Bayesian network} representing that model $\mathcal{M}$ is {\em transformable} into a model $\mathcal{M}'$ that produces a trajectory $\tau(\pi,P)$ that (1) reaches the goals in $P$ and (2) it is consistent with $\mathcal{O}(\tau)$.}
\label{fig:net}
\end{figure}

Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ likelihood is to assess the cost of transforming $\mathcal{M}$ into a model $\mathcal{M'}$ that produces a trajectory $\tau(\pi,P)$ such that: (1) $\tau(\pi,P)$ reaches the goals in $P$ and (2) $\tau(\pi,P)$ is consistent with observation $\mathcal{O}(\tau)$. Figure~\ref{fig:net} shows the {\em Bayesian network} that embodies this procedure. Regarding this network we have the following formulation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}:

\begin{align}
 P(\mathcal{O}|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\tau} P(\mathcal{M'}|\mathcal{M})P(\tau|\mathcal{M'})P(\mathcal{O}|\tau),
\end{align}

where $\tau$ ranges over all the trajectories consistent with $\mathcal{O}(\tau)$ that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.

The exact computation of $P(\mathcal{O}|\mathcal{M})$ with equation (2) is intractable. For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge, infinite in the case of planning problems without dead-ends~\cite{lesh1995sound}. Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially with the number of state variables. Instead, our approach is to estimate of $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} referred to \strips\ planning models. {\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, which have been proved successful for {\em pattern recognition}~\cite{MasekP80,Bunke97}. In this work, we assess the cost of {\em transforming} $\mathcal{M}$ into $\mathcal{M'}$ computing {\em edit distances} between \strips\ planning models.

\section{Recognition of \strips\ planning models}
\label{sec:asPlanning}

Our formalization of a model recognition task is valid for any planning model but our \emph{edit distance} measure is exclusively for \strips\ models and so we restrict our attention to \strips\texttt{-}compilable planning models\footnote{An edit distance for other planning models is also definable}. Thus, we focus on the recognition of $\mathcal{M}\in M$, where $M$ is a set of \strips\ planning models and $\mathcal{M}$ defines the semantics of the actions through a set of \strips\ action schemata.

\subsection{Well-defined \strips\ action schemata}
\strips\ action schemata provide a compact representation for specifying classical planning models. Figure~\ref{fig:model-example} shows six \strips\ action schemata that shape a particular kind of robot navigation in $N\times N$ grids (no matter the grid size).

{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.

@@

For instance, we claim that the six action schemata of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are not. Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.

Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$. The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$. We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$. For any of the six action schemata of Figure~\ref{fig:model-example}, the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.

@@

\begin{itemize}
\item {\em Syntactic constraints}. \strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$. Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$. For every action schema in the navigation model of Figure~\ref{fig:model-example} then $2^{2\times 10}=1,048,576$.

\item {\em Domain-specific constraints}. One can introduce domain-specific knowledge to constrain further the space of possible schemata. For instance, in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list. Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear in any of these lists because the {\tt\small next} predicate codes the {\em successor} function for {\em natural numbers}. These domain-specific constraints reduce further the size of the space of possible action schemata to $2^{2\times 7}=16,384$ (for every schema in the navigation model of Figure~\ref{fig:model-example}).
\end{itemize}

@@

\subsection{Edit distances for \strips\ planning models}
First, we define the two edit \emph{operations} on a schema $\xi$ that belongs to a \strips\ model $\mathcal{M}\in M$:

\begin{itemize}
\item {\em Deletion}. Given $\xi\in\mathcal{M}$, an element from any of the lists $pre(\xi)$/$del(\xi)$/$add(\xi)$ is removed such that the result is a {\em well-defined} \strips\ action schema.
\item {\em Insertion}. Given $\xi\in\mathcal{M}$, an element in ${\mathcal I}_{\Psi,\xi}$ is added to any of the lists $pre(\xi)$/$del(\xi)$/$add(\xi)$ such that the result is a {\em well-defined} action schema.
\end{itemize}

We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ models are. The distance is symmetric and meets the {\em metric axioms} provided that the two edit operations, {\em deletion} and {\em insertion}, have the same positive cost.

\begin{definition}[Edit distance]
  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} and {\em well-defined} \strips\ planning models within the same set of predicates $\Psi$. The {\bf edit distance} $\delta(\mathcal{M},\mathcal{M}')$ is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
\end{definition}

Since ${\mathcal I}_{\Psi,\xi}$ is a bounded set, the maximum number of edits that can be introduced to an action schema is bounded as well. The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built with predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that if we consider the set of syntactic constraints then $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).

\vspace{0.02cm}

An observation of the execution of a plan generated with $\mathcal{M}$ further constraints the space of possible action schemata of $\mathcal{M}$. The \emph{semantic knowledge} included in the observations introduce a third type of constraints, that we will call {\em observation constraints}, and that can be added to the set $\mathcal{C}$. In addition, {\em observation constraints} allow us to define an edit distance to elicit the value of $P(\mathcal{O}|\mathcal{M})$. It can be argued that the shorter this distance the better the given model explains the given observation.

\begin{definition}[Observation edit distance]
Given a planning problem $P$, an observation $\mathcal{O}(\tau)$ of the execution of a plan that solves $P$ and a \strips\ planning model $\mathcal{M}$ (all defined within the same set of predicates $\Psi$). The {\bf observation edit distance}, $\delta^o(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t. $\mathcal{M}'$ produces a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$; \[\delta^o(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
\end{definition}

$\delta^o(\mathcal{M},\mathcal{O})$ can also be defined through the edition that the observation $\mathcal{O}(\tau)$ requires to fit $\mathcal{M}$. This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of the model $\mathcal{M}$~\cite{yang2007learning,SohrabiRU16}. Our definition of {\em observation edit distance} is more practical since the size of ${\mathcal I}_{\Psi,\xi}$ is usually much smaller than $F$ (the number of variables in the action schemata should normally be lower than the number of objects in a planning problem).

\begin{definition}[{\em Closest consistent models}] \label{consistent}
Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2) their {\em edit distance} to $\mathcal{M}$ is minimal;
  \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
\end{definition}

\subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}

Now we we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ action schemata.

>>>>

%%%% ijcai19.tex

\typeout{IJCAI-19 Instructions for Authors}

% These are the instructions for authors for IJCAI-19.

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
% The file ijcai19.sty is NOT the same than previous years'
\usepackage{ijcai19}

% Use the postscript times font!
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}

%%%%%%%%%%%%%%%%%% Added for this paper
\usepackage{ wasysym }
\newcommand{\tup}[1]{{\langle #1 \rangle}}
\newcommand{\pre}{\mathsf{pre}}     % precondition
\newcommand{\del}{\mathsf{del}}     % effect
\newcommand{\add}{\mathsf{add}}     % effect
\newcommand{\eff}{\mathsf{eff}}     % effect
\newcommand{\cond}{\mathsf{cond}}   % conditional effect
\newcommand{\true}{\mathsf{true}}   % true
\newcommand{\false}{\mathsf{false}} % false
\newcommand{\PE}{\mathrm{PE}}     % precondition
\newcommand{\strips}{\textsc{Strips}}

@@

%%%%%%%%%%%%%55

% the following package is optional:
%\usepackage{latexsym}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

\title{Computing the least commitment action models from observations of plan executions}

% Single author syntax
%\author{
%    Sarit Kraus
%    \affiliations
%    Department of Computer Science, Bar-Ilan University, Israel \emails
%    pcchair@ijcai19.org
%}

% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
% Check the ijcai19-multiauthor.tex file for detailed instructions
\author{
Diego Aineto$^1$\and
Sergio Jim\'enez$^1$\and
Eva Onaindia$^1$\And
\and
Blai Bonet$^2$
\affiliations
$^1${\small Departamento de Sistemas Inform\'aticos y Computaci\'on. Universitat Polit\`ecnica de Val\`encia. Valencia, Spain}\\
$^2${\small Departamento de Computaci\'on. Universidad Sim\'on Bolívar. Caracas, Venezuela}
\emails
{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es, bonet@usb.ve}}

@@

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\section{Background}
This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
\subsection{Classical planning with conditional effects}
$F$ is the set of {\em fluents} or {\em state variables} (propositional variables). A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. $L$ is a set of literals that represents a partial assignment of values to fluents, and $\mathcal{L}(F)$ is the set of all literals sets on $F$, i.e.~all partial assignments of values to fluents. A {\em state} $s$ is a full assignment of values to fluents. We explicitly include negative literals $\neg f$ in states and so $|s|=|F|$ and the size of the state space is $2^{|F|}$.

@@

\subsection{Conformant planning}

\section{Learning the least commitment action models from observations}

\subsection{The learning task}

\subsection{The hypothesis space}
\strips\ action schemata provide a compact representation for specifying classical planning models. {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.

@@

For instance, the {\small\tt stack(?v1,?v2)} and {\small\tt unsstack(?v1,?v2)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} are not. Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.

Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$. The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$. We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.

@@

\begin{itemize}
\item {\em Syntactic constraints}. \strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$. Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.

\item {\em Domain-specific constraints}. One can introduce domain-specific knowledge to constrain further the space of possible schemata. For instance, in a {\em blocksworld} one can claim that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in any of these lists because of the semantic of the {\tt\small on} predicate that codes which block is in top of other block. 
\end{itemize}

@@

\subsection{A propositional encoding for \strips\ action schema}
Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$. Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.

The interest of having a propositional encoding for \strips\ action schema is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}. Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state. Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
\begin{small}  
\begin{align*}
\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
\cond(\mathsf{editable_{\xi}})=&\{del\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
&\{add\_e\_\xi\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
\end{align*}
\end{small}

Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}). Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.

\begin{figure}
  \begin{tiny}  
  \begin{verbatim}
(:action editable_inc-x
  :parameters (?v1 ?v2)
  :precondition
    (and (or (not (pre_xcoord_v1_inc-x)) (xcoord ?v1))
         (or (not (pre_xcoord_v2_inc-x)) (xcoord ?v2))
         (or (not (pre_ycoord_v1_inc-x)) (xcoord ?v1))                       
         (or (not (pre_ycoord_v2_inc-x)) (xcoord ?v2))
         (or (not (pre_q0__inc-x)) (q0))
         (or (not (pre_q1__inc-x)) (q1)))
         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2))))
    :effect (and
       (when (del_xcoord_v1_inc-x) (not (xcoord ?v1)))
       (when (del_xcoord_v2_inc-x) (not (xcoord ?v2)))
       (when (del_ycoord_v1_inc-x) (not (xcoord ?v1)))
       (when (del_ycoord_v2_inc-x) (not (xcoord ?v2)))
       (when (del_q0__inc-x) (not (q0)))
       (when (del_q1__inc-x) (not (q1)))
       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))

       (when (add_xcoord_v1_inc-x) (xcoord ?v1))
       (when (add_xcoord_v2_inc-x) (xcoord ?v2))
       (when (add_ycoord_v1_inc-x) (xcoord ?v1))
       (when (add_ycoord_v2_inc-x) (xcoord ?v2))
       (when (add_q0__inc-x) (q0))
       (when (add_q1__inc-x) (q1))
       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
  \end{verbatim}           
  \end{tiny}  
 \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
\label{fig:editable}
\end{figure}

\subsection{Learning the least commitment models with conformant planning}

\section{Evaluation}
\section{Conclusions}

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{planlearnbibliography.bib}
